Foreground Signatures and Mitigation in Neutral Hydrogen Power Spectra From The Reionization Epoch

### 2014-10-23 foreground_delay_spectrum_modeling.tex
from 4d54f99: ============
(*) Changed paragraph justifying discrepancy between data and simulations.
(*) Commented out the figure showing the histogram of log(rho) and instead
justified by text

In addition to intrinsic model uncertainty, delay spectra from simulations and data each have fluctuations due to thermal noise in the delay spectrum with rms $\sim$ 1.4$\times 10^7$~Jy~Hz. We estimate the ratio of delay spectra from data and simulations as $\rho = |V^\textrm{D}_{b\tau}(\vec{b},\tau)|\,/\,|V^\textrm{S}_{b\tau}(\vec{b},\tau)|$, where superscripts D and S denote data and simulation, respectively. % Using the aforementioned uncertainties added in quadrature, we estimate the approximate expected uncertainty in $\log_{10}\rho$ and denote it by $\Delta\,\log_{10}\rho$.

Figure~\ref{fig:data-sim-ratio} shows normalized histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}. The median absolute deviation from these distributions corresponds to $\sim$~90\% fractional difference between data and modeling on average in either case. This is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account. Currently, we are significantly limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
\caption{Normalized histograms of logarithm of ratios of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Median absolute deviation of $\log_{10}\rho$ is $\sim 0.28$ in both cases indicating a fractional difference of $\sim 90$\%. A value of zero indicates perfect agreement ($\rho=1$). \label{fig:data-sim-ratio}}
\end{figure}

These uncertainties are presented to confirm the qualitative agreement already seen between Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}. These uncertainty estimates are conservative. A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power--pattern \citep{ber14}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations, but is beyond the scope of this paper. Our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.

@@

In subsequent sections, we provide a detailed explanation of our results as a combination of factors noted above. Note that numerous features may overlap at different degrees of significance depending on combinations of parameters. We assign the features to their predominant causes. Secondly, we have used noiseless cases to clearly illustrate the observed foreground signatures. With the addition of noise in the visibilities which is subject to observing time, some of the weaker features may not be as prominently visible. Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as antenna spacing length and orientation, power--pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectra of foreground emission.

>>>>

In addition to intrinsic model uncertainty, delay spectra from simulations and data each have fluctuations due to thermal noise in the delay spectrum with rms $\sim$ 1.4$\times 10^7$~Jy~Hz. We estimate the ratio of delay spectra from data and simulations as $\rho = |V^\textrm{D}_{b\tau}(\vec{b},\tau)|\,/\,|V^\textrm{S}_{b\tau}(\vec{b},\tau)|$, where superscripts D and S denote data and simulation, respectively.

The median absolute deviation of $\log_{10}\rho$ inside the {\it foreground wedge} for both pointings is $\approx 0.28$. This corresponds to $\sim$~90\% fractional difference between data and modeling on average with either pointing. We also simulated delay spectra after assigning spectral indices drawn randomly from a {\it gaussian} distribution with a mean of $\langle\alpha_\textrm{sp}\rangle=-0.83$ and a HWHM of $\Delta\alpha_\textrm{sp}=0.35$ to the point sources in our compact foreground model. These simulations typically yielded a median absolute deviation of $\approx 0.29$ for $\log_{10}\rho$ indicating fractional differences of $\sim$~95\%. This demonstrates that a fractional deviation of $\sim$~90\% observed between data and simulations is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account. Currently, we are significantly limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.

% Figure~\ref{fig:data-sim-ratio} shows normalized histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}. The median absolute deviation from these distributions corresponds to $\sim$~90\% fractional difference between data and modeling on average in either case. This is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account. Currently, we are significantly limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.

% \begin{figure}[htb]
% \centering
% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
% \caption{Normalized histograms of logarithm of ratios of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Median absolute deviation of $\log_{10}\rho$ is $\sim 0.28$ in both cases indicating a fractional difference of $\sim 90$\%. A value of zero indicates perfect agreement ($\rho=1$). \label{fig:data-sim-ratio}}
% \end{figure}

These uncertainties are presented only to confirm the qualitative agreement already seen between Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}. These estimates are conservative. A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power--pattern \citep{ber14}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations, but is beyond the scope of this paper. Our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.

@@

In subsequent sections, we provide a detailed explanation of our results as a combination of factors noted above. Note that numerous features may overlap at different degrees of significance depending on combinations of parameters. We assign the features to their predominant causes. Secondly, we have used noiseless cases to clearly illustrate the observed foreground signatures. With the addition of noise in the visibilities, some of the weaker features may not be as prominently visible. Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as antenna spacing length and orientation, power--pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectra of foreground emission.
